---
title: "A conclusive title statement: An instructional guide for generating powerpoint files from Rmarkdown files for enabling productive scientific communication at BillionToOne"
author: David Tsao
date: "May 16, 2024"
footer: "BillionToOne, Inc."
output:
  powerpoint_presentation:
    reference_doc: ./bto_template/BTO-template.potx
---

```{r echo=F}
# Setup chunk options. echo = False prevents R code from being printed
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)
knitr::opts_chunk$set(fig.width=8, fig.height=4)

```

```{r}
# import packages
suppressPackageStartupMessages({library(tidyverse)})
```

# Introduction & Motivation

## Science is generating new knowledge and telling people about it

- If a tree falls in a forest and no one is around to hear it, does it make a sound?
- Does art have any value if no one is around to view it?
- Do your scientific results have any worth if no one understands it?

## Claim: Generating powerpoint slides directly from RMarkdown can increase scientific communication

The technical report format is great for communicating scientific results. However, the "scientific" talk is also an important scientific communication method.

For external talks, the level of polish and background setting needs to be very high. It can take 10+ hours to create a new talk for external audiences from scratch.

Internal talks (e.g. group meeting) are more informal, and the audience has much more context. 

Herein, I show how to use Rmarkdown to efficiently create scientific talks for internal purposes. 

# Methods


## Rmarkdown to powerpoint pipeline

A modified "build" pipeline was used to generate Powerpoint presentations using Rmarkdown / Rstudio IDE / pandoc.
The pipeline was adapted from the posit solutions engineering [powerpoint vignette](https://github.com/sol-eng/powerpoint). 

Briefly, a custom powerpoint template was created, BTO-template.potx, that has BillionToOne brand colors and logos. Additionally, presentation guidelines were followed from https://www.arl.org/accessibility-guidelines-for-powerpoint-presentations/, such as minimum 24pt font size.

As a proof of concept, this presentation was created using the template and Rmarkdown. Examples were included for sections, bullets, plotting, and tables.

# Results

## Text formatting can be efficiently applied using markdown syntax

:::::::::::::: {.columns}
::: {.column}
```
*italic* 
**bold** 
~~strikeout~~ 
~subscript~
^superscript^
[small caps]{.smallcaps} 
`verbatim`
```
:::
::: {.column}
*italic* 

**bold** 

~~strikeout~~ 

~sub~script

^super^script

[small caps]{.smallcaps} 

`verbatim`
:::
::::::::::::::

## Latex can be used for equations

Here is an example. Show the Euler identity:

\[ e^{i\pi} = -1  \]


Begin by Taylor expanding $e^{i\theta}$:

\[e^{i\theta} = 1 + i\theta + (i\theta)^2/2 + (i\theta)^3/3! +  (i\theta)^4/4! + ...\]

Notice that the odd and even terms are just the taylor expansion of $\cos$ and $\sin$:

\[e^{i\theta} = 1 - \theta^2/2 + \theta^4/4! +  i(\theta - \theta^3/3!) + ... \]

\[ e^{i\theta} = \cos(\theta) +  i\sin(\theta) \]

## Plots and legends are easily embedded with text

### Diamonds dataset

The price of a diamond increases super-linearly with the carat of the diamond. 

```{r caption = "Canonical diamonds data visualization", fig.width=8, fig.height=4}
ggplot(diamonds) + 
  geom_point(aes(x = carat, y = price, color = color, shape = cut))
```

## How to show a table

Tables are shown using `knitr::kable`. 
```{r}
diamonds %>%
head() %>%
knitr::kable()
```

## 1-column plots or tables

```{r}
# Don't include anything except for the code block
diamonds %>%
head() %>%
knitr::kable()
```


# Discussion & Conclusions

- Talks are an effective method of scientific communication
- Creating talks takes a long time
- Powerpoints can be quickly created through integration with Rstudio / Rmarkdown can 